\section{Proofs and Scope of the Curvature Analysis}
\label{app:theory-proofs}

\subsection{Taylor expansion and the role of a PSD proxy}

Let $\mathcal{L}:\mathbb{R}^p\rightarrow\mathbb{R}$ be twice
continuously differentiable on the segment from $w_0$ to
$w_0+\delta$.  With $J(w)=\nabla^2\mathcal{L}(w)$, Taylor's integral
form gives
\begin{equation}
\label{eq:integral-taylor}
\mathcal{L}(w_0+\delta)-\mathcal{L}(w_0)
=g^\top\delta+
\int_0^1(1-t)\,\delta^\top J(w_0+t\delta)\delta\,dt.
\end{equation}
Writing $J_0=J(w_0)$ yields the usual quadratic term and a path
remainder.  If
$\|J(w_0+t\delta)-J_0\|_{\mathrm{op}}\le Mt\|\delta\|_2$, then
\begin{equation}
\left|\mathcal{L}(w_0+\delta)-\mathcal{L}(w_0)
-g^\top\delta-\tfrac12\delta^\top J_0\delta\right|
\le \frac{M}{6}\|\delta\|_2^3,
\end{equation}
because $\int_0^1t(1-t)dt=1/6$.

The task Hessian $J_0$ can be indefinite.  For geometric diagnostics we
instead select and disclose $H\succeq0$.  Combining proxy mismatch and
path variation gives
\begin{equation}
\label{eq:proxy-error-bound}
\begin{split}
\varepsilon_H(\delta)
={}&\tfrac12\delta^\top(J_0-H)\delta+R_3(\delta),\\
|\varepsilon_H(\delta)|
\le{}&\tfrac12\|J_0-H\|_{\mathrm{op}}\|\delta\|_2^2\\
&+\tfrac{M}{6}\|\delta\|_2^3.
\end{split}
\end{equation}
The bound need not be numerically tight; its purpose is to expose the
assumption hidden when a PSD surrogate is called ``the Hessian.''  For
a damped task Hessian, $H=J_0+\lambda I$ is a metric only when
$\lambda\ge-\lambda_{\min}(J_0)$.  Damping must therefore be reported,
and it changes the measured angles.  If $H$ is singular, its inner
product is understood on the quotient by $\ker(H)$; correlations with
$q_i=0$ are undefined rather than zero.

For sequential maps $T_1,\ldots,T_k$, let
$w_i=T_i(w_{i-1})$ and $\delta_i=w_i-w_{i-1}$.  Then
$w_k-w_0=\sum_i\delta_i$, so substituting into the $H$-quadratic gives
\begin{equation}
\frac12\Bigl\|\sum_i\delta_i\Bigr\|_H^2
=\frac12\sum_i\|\delta_i\|_H^2
+\sum_{i<j}\langle\delta_i,\delta_j\rangle_H.
\end{equation}
This algebra does not make the components order invariant.  If
$T_iT_j(w)\ne T_jT_i(w)$, changing order changes both increments and
the final endpoint; all terms must be recomputed after final decoding.

\subsection{Coherence bound}

\paragraph{Proposition 1 (PSD coherence).}
Let $H\succeq0$, let $q_i=\|\delta_i\|_H^2>0$, and suppose
$|\rho_H(i,j)|\le\mu$ for every $i\ne j$.  Then
Equation~\eqref{eq:coherence-bound} holds.

\paragraph{Proof.}
Cauchy--Schwarz in the PSD seminorm gives
$|\langle\delta_i,\delta_j\rangle_H|
\le\mu\sqrt{q_iq_j}$.  Moreover,
\begin{equation}
2\sum_{i<j}\sqrt{q_iq_j}
=\left(\sum_i\sqrt{q_i}\right)^2-\sum_iq_i
\le(k-1)\sum_iq_i,
\end{equation}
where the last inequality follows from
$(\sum_i\sqrt{q_i})^2\le k\sum_iq_i$.  Applying the upper and lower
signs to the cross-term expansion proves the claim. \hfill$\square$

The proposition controls only additivity of the PSD quadratic.  It
does not control $g^\top\delta$, proxy error, path remainder, component
self costs, or serialized rate.  A negative correlation may reduce the
quadratic through cancellation, but $|\rho_H|$ is then large and the
components are not near-orthogonal.

\subsection{Feasibility of enforced orthogonality}

Suppose an error $e$ can be modified using columns
$B=[b_1,\ldots,b_d]$ and we want the result to be $H$-orthogonal to
columns $U=[u_1,\ldots,u_m]$.  A coefficient vector $\alpha$ must solve
\begin{equation}
\label{eq:orthogonal-feasibility}
A\alpha=-c,\qquad A=U^\top HB,\qquad c=U^\top He.
\end{equation}
An unconstrained solution exists if and only if $-c$ lies in the range
of $A$.  Thus a single global scale generally cannot remove several
independent moments.  Codec feasibility adds stronger conditions:
$\alpha$ must respect the allowed scale range, support or rank, storage
precision, and descriptor layout.  After rounding, the residual
$\|A\widehat\alpha+c\|$ must be remeasured.  Subsequent requantization,
support selection, or rank truncation can reintroduce the moments, so
orthogonality is a property of the final decoded endpoint, not of an
intermediate floating-point solve.

\subsection{A sufficient endpoint comparison}

Define the local PSD score
\begin{equation}
\widehat\Delta_H(z)=g^\top z+\tfrac12 z^\top Hz
\end{equation}
and suppose a candidate-specific certificate satisfies
\begin{equation}
\left|\mathcal{L}(w_0+z)-\mathcal{L}(w_0)
-\widehat\Delta_H(z)\right|\le r(z).
\end{equation}
The radius $r(z)$ must include proxy mismatch and path curvature; it may
be estimated conservatively from held-out interpolation diagnostics but
is not supplied by coherence alone.

\paragraph{Proposition 2 (same-byte sufficient condition).}
Let $z_C$ be a composite candidate and $\{z_j\}$ the eligible
single-mechanism candidates.  If
\begin{equation}
\label{eq:certified-comparison}
\begin{split}
\mathcal{B}(z_C)&\le B,\\
\widehat\Delta_H(z_C)+r(z_C)
&<\min_{j:\,\mathcal{B}(z_j)\le B}
  \bigl\{\widehat\Delta_H(z_j)-r(z_j)\bigr\}.
\end{split}
\end{equation}
then the composite has lower loss than every eligible single candidate
at budget $B$.

\paragraph{Proof.}
The left side of the strict inequality upper-bounds the composite loss
increment, while every term on the right lower-bounds the corresponding
single-candidate increment. \hfill$\square$

This proposition is deliberately sufficient, not necessary.  In the
absence of a trustworthy $r(z)$, held-out endpoint loss replaces the
certificate; a high proxy/path correlation is a diagnostic, not a
proof of the inequality.

\subsection{Comfort-zone allocation}

Let each $d_i:[0,S]\rightarrow\mathbb{R}$ be differentiable, convex,
and nondecreasing in assigned compression burden.  Consider
\begin{equation}
\label{eq:continuous-allocation}
\min_{s_i\ge0}\sum_{i=1}^k d_i(s_i)
\quad\text{subject to}\quad \sum_{i=1}^k s_i=S.
\end{equation}

\paragraph{Proposition 3 (marginal-cost allocation).}
At an optimum of Equation~\eqref{eq:continuous-allocation}, every
active component satisfies $d_i'(s_i)=\lambda$ and every inactive
component satisfies $d_i'(0^+)\ge\lambda$.  If the all-$A$ allocation
has $d_A'(S)>d_B'(0^+)$ and both derivatives are continuous near the
endpoint, then an allocation using both $A$ and $B$ has strictly lower
cost.

\paragraph{Proof.}
The KKT stationarity and complementary-slackness conditions are
\begin{equation*}
d_i'(s_i)-\lambda-\nu_i=0,\qquad
\nu_i\ge0,\qquad \nu_i s_i=0.
\end{equation*}
They give the stated active and inactive conditions.  For the second
claim, set $f(\epsilon)=d_A(S-\epsilon)+d_B(\epsilon)$.  Then
$f'(0^+)=-d_A'(S)+d_B'(0^+)<0$, so sufficiently small positive
$\epsilon$ improves the objective. \hfill$\square$

Separability is an assumption of Equation~\eqref{eq:continuous-allocation}.
For an objective $\sum_i d_i(s_i)+\gamma(s)$ with non-negligible
allocation-dependent interactions, the active condition becomes
$d_i'(s_i)+\partial_i\gamma(s)=\lambda$.

The proposition explains why several mechanisms can outperform one
mechanism pushed beyond its comfortable region, but only in the stated
continuous model.  A codec has a lattice of feasible endpoints: one
rank, one sparse group, or a new descriptor may have a fixed charge.
The deployable analogue is a finite exact-byte exchange.

Let $A$ be a realized endpoint, $u$ a perturbation that frees payload,
and $r$ a repair perturbation.  For a fixed, order-specific decomposition
$C=A+u+r$, direct addition and subtraction gives
\begin{equation}
\label{eq:exchange-identity}
\mathcal{L}(C)-\mathcal{L}(A)=P_A(u)-G_A(r)+\Gamma_A(u,r).
\end{equation}
Hence Equation~\eqref{eq:exact-exchange} is equivalent to a strict
improvement once $\mathcal{B}(C)\le\mathcal{B}(A)$.  Around $A$, the
second-order part of $\Gamma_A$ is $u^\top J_A r$; replacing $J_A$ by a
PSD proxy again adds a mismatch and path term.  This identity also shows
why repair-to-repair orthogonality is insufficient: the cost of freeing
bytes and all interactions still remain.

\subsection{Physical bytes and parameter utilization}

For endpoint $z$, the comparison rate is the on-disk length
\begin{equation}
\label{eq:byte-ledger}
\begin{split}
\mathcal{B}(z)={}&B_{\rm header}+B_{\rm manifest}+B_{\rm descriptor}\\
&+B_{\rm codes}+B_{\rm scales}\\
&+B_{\rm sparse\ values}+B_{\rm indices}\\
&+B_{\rm rowptr}+B_{\rm factors}+B_{\rm alignment}\\
&+B_{\rm tail\ padding}.
\end{split}
\end{equation}
The marginal cost $m_r=\mathcal{B}(A+r)-\mathcal{B}(A)$ is therefore
integer-valued and need not equal the sum of value streams.  Alignment
can make it nonadditive.  We disclose both natural bytes and final file
bytes: the latter decides the strict budget, while the former reveals
when a nominally zero-cost change merely consumes padding slack.

When $m_r>0$, held-out recovery per byte is
\begin{equation}
\eta_{\rm NLL}(r\mid A)
=\frac{\mathcal{L}(A)-\mathcal{L}(A+r)}{m_r}.
\end{equation}
Suppose $u$ frees $f_u>0$ bytes and has loss penalty
$P_A(u)=f_u\eta_{\rm giveup}$.  If the repair has
$G_A(r)=m_r\eta_{\rm repair}$, then a feasible exchange improves loss
when
\begin{equation}
m_r\eta_{\rm repair}
>f_u\eta_{\rm giveup}+\Gamma_A(u,r),
\qquad m_r\le f_u,
\end{equation}
with any remaining slack represented in the actual serialization.
For an exactly matched exchange $m_r=f_u$, this reduces to an efficiency
comparison corrected by interaction per byte.  Merely comparing the two
ratios is not sufficient when the exchanged byte amounts differ.

If $m_r=0$, division would create an artificial infinite efficiency.
Instead we report the direct recovery, verify that the decoded artifact
contains no side state, and state whether natural bytes or only final
file bytes are unchanged.  Fixed-support OBS can satisfy this condition
when re-estimated values replace already stored values; a scale can do
so only if it is demonstrably folded into an already stored field.

\subsection{Limits of the diagnostic}

The analysis is local and order specific.  Its strongest threats are:
(i) noncommuting quantization, support selection, and low-rank fitting;
(ii) curvature drift between $w_0$ and the decoded endpoint;
(iii) an activation/Fisher/GGN proxy that does not preserve task-loss
ranking; (iv) calibration geometry that does not transfer to held-out
text, seeds, layers, or model families; and (v) discrete metadata and
alignment thresholds.  These threats motivate final-endpoint NLL,
multiple independent splits, source and artifact hashes, and full byte
round trips.  The strongest fine-grained numerical evidence remains one seed
and six selected MLP linear tensors of Pythia-70M.  Three additional, unpooled
seed-17 scalability-smoke jobs cover full Pythia-70M MLP scope and
depth-stratified OPT-125M and Qwen3-0.6B MLP scopes at one selected-weight
artifact rate.  The propositions and smoke observations do not upgrade these
probes into a multi-seed, whole-model, or cross-model advantage claim.
