\section{Experimental Protocol}
\label{sec:experiments}

\paragraph{Verified exploratory endpoint.}
The currently verified artifact uses the pinned Pythia-70M checkpoint
\citep{biderman2023pythia}, WikiText-2 raw validation text
\citep{merity2017wikitext}, and six MLP projections from the first,
middle, and last transformer blocks.  Calibration and evaluation source texts
are content-disjoint; fallback text is disabled.  Held-out metrics cover
\VerifiedEvalTokens{} next-token predictions and dense perplexity
\DensePPL{}.  The target selected-tensor artifact ratio is 0.258 relative to
an FP16 reference.  Six strategies are probed on 13-point paths, and 16 fixed
contiguous windows provide paired descriptive differences.  These windows are
not treated as independent population samples.

\paragraph{Completed scalability smoke.}
We execute every native endpoint once at seed \ScalePilotSeed{} and target
selected-weight artifact ratio 0.258 on three scopes.  Pythia-70M uses all
\ScalePilotPythiaSelectedTensors{} transformer-block MLP projections;
OPT-125M~\citep{zhang2022opt} uses \texttt{fc1/fc2} in five depth-stratified
blocks (\ScalePilotOPTSelectedTensors{} tensors); and
Qwen3-0.6B~\citep{yang2025qwen3} uses \texttt{up/down} projections
in five blocks, excluding its gate projection
(\ScalePilotQwenSelectedTensors{} tensors).  Every job evaluates
\ScalePilotEvalTokens{} next-token predictions in eight fixed windows and
probes six directions on the same 13-point path.  All three share numerical
source \ScalePilotNumericalSourceSHA{}, record a per-layer PSD covariance
audit, and pass round-trip SHA and exact-file-byte checks.  The jobs are three
separate observations: each uses WikiText-2 raw validation text with
content-disjoint calibration/evaluation source windows and fallback disabled;
tokenizers and tensor scopes differ, and no cross-model metric is pooled or
ranked.

\paragraph{Preregistered and planned scale matrix.}
The released 69-job configuration also declares Pythia-160M/410M and
Llama-3.2-1B~\citep{grattafiori2024llama3}, multiple rates, and additional
seeds.  Those rows remain unexecuted.  Moreover, the present runner's seed
argument does not alter its sequential data windows, so repeated invocations
cannot be treated as independent confirmation.  The separate eight-seed
content-disjoint split manifest is preregistered but not yet consumed by this
runner.

\paragraph{Baselines and fairness.}
Native comparisons include Q, global and block scales, Q+S, Q+S with OBS,
Q+L, strict and relaxed QSL, Q+S$_{\rm OBS}$+L, and component-scaled QSL.
External methods enter a numeric table only if model checkpoint, W-only scope,
tensor manifest, calibration/evaluation data, and exported physical payload
match.  We report data-free A0, calibrated no-backward A1, local
backward-assisted B, and global fine-tuned/QAT C methods separately
(\cref{tab:lane-summary,tab:lanes}).  Literature numbers never populate
measured cells.

\paragraph{Metrics.}
Primary metrics are final file bytes, held-out NLL and perplexity change,
normalized PSD proxy cost, and all pairwise signed $\rho_H$.  For a positive
byte increment we use held-out NLL recovery per added byte.  Equal-byte
re-estimation is reported as direct recovery, not infinite efficiency.
Natural bytes, alignment, encoder wall time, peak memory, and provenance are
secondary diagnostics.
