\section{Introduction}
\label{sec:introduction}

Weight-only post-training compression is no longer a single-knob problem.
Second-order quantizers, activation-aware scales, one-shot pruning, sparse
outliers, and low-rank error reconstruction all preserve different parts of a
pretrained model~\citep{frantar2023gptq,frantar2023sparsegpt,lin2024awq,dettmers2024spqr,zhang2024lqer}.
This diversity motivates a natural hypothesis: if every mechanism has a
small-perturbation regime in which loss grows slowly, then splitting a severe
compression target across several mechanisms could remain inside several
comfort zones and outperform concentrating all repair bytes in one mechanism
over the same mandatory quantized base.

Joint mechanisms and their interactions are not new by themselves.  SLiM
combines quantization, semi-structured sparsity, and low-rank compensation;
Harma et al. analyze Q/S order and non-orthogonality; OBR applies Hessian
compensation to joint quantization and sparsification; and prior rate--distortion
work couples Hessian or Fisher sensitivity to entropy or bit allocation
~\citep{mozaffari2025slim,harma2025effective,guo2026optimal,choi2017towards,orr2025optimal,young2025radio}.
We therefore study a narrower reproducibility question: what conclusions
survive when realized heterogeneous endpoints are decoded from a declared
self-describing artifact and compared by natural and final file bytes?

Two gaps obstruct a clean test.  First, a nominal bit rate does not determine a
heterogeneous artifact's size.  Sparse corrections need support indices and row
pointers; low-rank corrections need two factors; scales and codebooks need
descriptors; containers need headers and alignment.  A value-stream calculation
can therefore call two endpoints equal-rate even when their decoded files
differ.  Second, a small Hessian-weighted perturbation is not itself an
accuracy certificate.  The task Hessian may be indefinite, a PSD proxy may not
match it, the first-order term need not vanish, and curvature can drift along
the path to the final codec endpoint.

We study these gaps through \emph{exact-byte Hessian diagnostics}.  A
deterministic research codec serializes the selected tensors, round-trips every
endpoint, and charges codes, scales, sparse state, factors, descriptors, and
padding.  In a declared PSD activation-Gram geometry, realized Q, S, and L
increments are decomposed into self and signed cross terms.  A 13-point
interpolation then checks whether the local proxy tracks held-out NLL all the
way to the only deployable point, $\epsilon=1$.

Our contributions are:
\begin{itemize}
  \item an executable, byte-faithful protocol and fail-closed artifact
  validator, exact under the declared research codec, for
  quantization (Q), scale protection, sparse residuals (S), fixed-support
  Optimal Brain Surgeon (OBS)
  value repair, low-rank residuals (L), and their compositions;
  \item a theory that states what PSD near-orthogonality does guarantee and a
  stricter same-byte exchange condition that includes gradient, interactions,
  path error, and discrete byte overhead;
  \item a parameter-utilization analysis based on held-out loss recovered per
  added file byte, with zero-byte repairs reported separately; and
  \item an evidence-aware comparison framework separating data-free PTQ,
  calibrated frozen PTQ, backward-assisted correction, and global
  fine-tuning/quantization-aware training (QAT).
\end{itemize}

The current strict result is intentionally negative.  On one \VerifiedModel{}
run over \VerifiedTensorCount{} selected MLP tensors, Q+L and scaled Q+S+L are
both \QLBytes{} bytes, but their perplexity increases are \QLDeltaPPL{} and
\StrictQSLDeltaPPL{}.  This verified physical endpoint shows that a logical
payload estimate is insufficient for the operational comparison.  Its strict
QSL natural artifact is \StrictQSLTailPaddingBytes{} bytes below Q+L and is
tail-padded to equal final length, so this result rejects the tested
conservative per-layer candidate rather than a budget-exhausted global
frontier.  It does not establish that composite compression is universally
inferior.  We additionally
execute all native endpoints in three separate single-seed scalability-smoke
jobs: full Pythia-70M MLP scope and depth-stratified OPT-125M and Qwen3-0.6B
MLP scopes.  All three retain a positive strict-scaled-QSL-minus-QL NLL sign
despite near-orthogonal S--L proxy geometry.  This repeated diagnostic enlarges
the execution scope, but unequal tensor scopes and one seed preclude a pooled
or model-scale conclusion.
