\section{Exact-Byte Composite Compression}
\label{sec:method}

\paragraph{Scope and objective.}
Let $\mathcal{T}$ be an ordered manifest of eligible weight matrices and
$w_0$ their dense FP16 values.  A codec endpoint $z$ decodes to
$\widehat w(z)$ and has on-disk length $\mathcal{B}(z)$.  For a fixed
manifest, model, and held-out distribution, the discrete problem is
\begin{equation}
  \min_{z\in\mathcal{Z}}\;
  \mathcal{L}(\widehat w(z))
  \quad\text{s.t.}\quad \mathcal{B}(z)\le B.
  \label{eq:discrete-rate}
\end{equation}
The same $\mathcal{T}$ is used for every comparator; all other model weights
remain dense.  Thus bytes in this paper describe the selected-tensor artifact,
not a whole checkpoint or inference package.

\paragraph{Byte-faithful codec.}
For each tensor, the artifact stores packed signed Q codes, FP16 row or
row--column-block scales, optional FP16 sparse values with fixed-width
compressed sparse row (CSR)
support, and optional FP16 low-rank factors.  A deterministic manifest stores
shapes, dtypes, stream offsets, and alignment.  We report both natural bytes
and final file bytes after 64-byte alignment and tail padding, then verify file
length, SHA-256, decoding, and elementwise equality with the evaluated FP16
endpoint.  This is an auditable research format, not a claim of production
kernel support.

\paragraph{Candidate mechanisms.}
Q is symmetric row-wise 4-bit round-to-nearest quantization.  Block-scale
protection solves bounded row--column-block multipliers in the local PSD
metric and stores the resulting FP16 scales.  S selects a residual support by
Wanda-style activation--magnitude importance; a matched control stores the
naive residual values.  OBS then re-estimates values on the \emph{same}
support, replacing already-paid values without changing the artifact layout.
L applies a covariance-whitened SVD to the post-Q residual and stores two FP16
factors.  Component scales for Q/S/L are solved jointly and folded into their
existing FP16 state, so a zero-byte claim is accepted only after natural and
file-byte verification.  The verified legacy endpoints use the declared
Q$\rightarrow$S$\rightarrow$L construction order; the new unexecuted matrix
also exposes L$\rightarrow$S as an order ablation.  We call Q+L a
\emph{single repair over the mandatory Q
base}, and Q+S+L a composite repair; neither phrase means that Q is absent.

\paragraph{Implemented PSD proxy.}
For layer $j$, calibration activations define a covariance $C_j$ and
\begin{equation}
 \begin{aligned}
 q_j(\Delta W)
   &=\tfrac12\operatorname{tr}(\Delta W C_j\Delta W^\top),\\
 \rho_H(a,b)
   &=\frac{\sum_j\langle a_j,b_j\rangle_{C_j}}
   {\sqrt{\sum_j\|a_j\|_{C_j}^2\sum_j\|b_j\|_{C_j}^2}}.
 \end{aligned}
 \label{eq:implemented-proxy}
\end{equation}
The reported normalized cost is
$\sum_j\|\Delta W_j\|_{C_j}^2/
 \sum_j\|W_j\|_{C_j}^2$.  Each empirical Gram receives a
$10^{-5}$ mean-diagonal ridge.  Matrices with relative negative spectrum worse
than $10^{-7}$ fail; accepted float32 numerical error is repaired once, and an
eight-machine-epsilon target and the applied diagonal shift are recorded per
layer.  Damping and the repair are identity shifts, so the final spectral
lower bound is derived from one validated eigenspectrum instead of repeating
cubic audit decompositions.  Whitened fitting retains one separate cached
eigendecomposition used to construct its square-root maps.  The same
prepared $C_j$ is used by selection, OBS, endpoint costs, Gram terms, and
$\rho_H$.  Whitened SVD has a separately disclosed numerical fit metric
$\widetilde C_j$: eigenvalues below a configurable fraction of
$\lambda_{\max}(C_j)$ are floored before forming the square root and inverse
square root.  The runner records that fraction, the number of clipped
eigenvalues, and the resulting condition number per layer.  Every decoded
candidate is still scored in $C_j$; factorizer regularization is not silently
presented as the common scoring metric.
For the unexecuted large-model smoke, each residual fit uses a seeded
randomized SVD with four oversampling vectors and two power iterations; the
resolved solver is recorded per layer.  Full SVD remains an explicit
small-matrix control, so computational approximation is not hidden inside the
geometry ablation.

\paragraph{Discrete strict-rate selection.}
For each layer and target, we enumerate feasible low-rank ranks.  At each rank,
the largest CSR support fitting the requested serialized cap is constructed;
the candidate with minimum activation-Gram cost is retained.  The strict QSL
control uses that layer's Q+L artifact as its cap, reserves alignment slack,
and is rejected before NLL evaluation if the aggregate serialized file exceeds
the Q+L file.  This is a deterministic layerwise constrained search, not a
claim of global discrete optimality.  It yields an executable endpoint rather
than a fractional rank or value-only relaxation.  A relaxed QSL point is
reported separately to show the effect of the next discrete allocation step.
The new, unexecuted aggregate allocator removes the forced-composition condition:
each layer may select Q, Q+S$_{\rm OBS}$, Q+L, or an enumerated QSL state.  It
adds intermediate supports and 1.25/1.5/2.0 multiples of the local Q+L repair
allowance beyond base Q near the strongest rank states, thereby permitting
enumerated cross-layer byte borrowing.  It prunes an additive one-layer
byte/cost surrogate and checks every surviving combination with the exact
multi-layer serializer.  The additive pruning is heuristic; feasibility is
exact under the research codec, while optimality remains limited to this
enumerated budget-band state set.  We therefore do not call it a globally
optimal frontier.  Component scaling is fitted only after the unscaled QSL
allocation is selected, so the scaled row is a fixed-allocation post-selection
ablation rather than the optimum scaled candidate over the whole frontier.

\paragraph{Endpoint gate.}
For every retained endpoint, we evaluate
$w(\epsilon)=w_0+\epsilon(\widehat w-w_0)$ on held-out text at 13 fixed
values from 0 to 1.  The first five points through $\epsilon=1/8$ fit a local
linear--quadratic model; the full path measures proxy/NLL correlation.  Only
$\epsilon=1$ is serialized and eligible for a rate--accuracy conclusion.
