\section{Related Work and Comparison Lanes}
\label{sec:related}

\paragraph{Frozen post-training methods.}
GPTQ uses second-order information to update remaining weights during
quantization, while SparseGPT applies an analogous one-shot view to
pruning~\citep{frantar2023gptq,frantar2023sparsegpt}.  Wanda replaces the
second-order solve with an activation--magnitude score~\citep{sun2024wanda}.
AWQ searches activation-aware scales that protect salient
channels~\citep{lin2024awq}.  SpQR and SqueezeLLM combine dense low-bit state
with sparse high-precision state, but the latter's Fisher-sensitivity path
requires loss gradients and therefore belongs in a different compute
lane~\citep{dettmers2024spqr,kim2024squeezellm}.  LQER, QERA, EoRA, and SLiM
reconstruct compression error in low-rank or joint sparse/quantized
representations~\citep{zhang2024lqer,zhang2025qera,liu2024eora,mozaffari2025slim}.
OBR applies closed-form Hessian compensation to joint quantization and
sparsification~\citep{guo2026optimal}.  Optimal Brain Surgeon and Optimal
Brain Compression already establish the broader second-order correction
lineage~\citep{hassibi1993obs,frantar2022obc}.  Fixed-support OBS refitting,
low-rank reconstruction, and Q/S/L composition are therefore baselines and
mechanisms in this paper, not base novelty claims.

\paragraph{Interaction, orthogonality, and rate precedents.}
Harma et al. prove that magnitude sparsification and max-scaled quantization
are generally non-orthogonal and analyze their order-dependent compound
error~\citep{harma2025effective}.  Their transformation-level notion differs
from our signed, order-specific cross term under a common local PSD proxy, but
it precludes any claim that this paper is the first interaction or
orthogonality analysis.  ProjQ likewise coordinates quantization noise with a
low-rank adapter through orthogonal subspace projection~\citep{yu2026projq}.
Choi et al. combine Hessian-weighted distortion with Huffman coding and
entropy-constrained scalar quantization~\citep{choi2017towards}; Orr et al.
study variable-length weight formats and Fisher rate allocation
~\citep{orr2025optimal}; and Radio optimizes an LLM rate--distortion
objective~\citep{young2025radio}.  Our narrower target is an executable,
byte-faithful audit under a declared research codec: every compared endpoint
is round-trip decoded, its natural and padded file bytes are separated, and
its signed local-PSD interactions are checked against realized NLL.  This is
not a claim of the first Hessian-aware rate allocation or the first compressed
bitstream.

\paragraph{End-loss, adaptive, and trained recovery.}
GuidedQuant uses end-loss information and YAQA performs full-model adaptive
rounding; these works make
the evaluation objective more explicit but use different optimization
budgets~\citep{kim2025guidedquant,tseng2026yaqa}.
Q-Palette targets fractional-bit allocation with codebooks~\citep{lee2025qpalette}.
LiftQuant adds block correction and an optional end-to-end stage, while
FPTQuant learns function-preserving transformations~\citep{he2026liftquant,vanbreugel2026fptquant}.
SliderQuant learns fused channel scales and small low-rank auxiliary repairs
under local sliding-window reconstruction, making it a close lane-B analogue
to our scale-plus-low-rank parameter-utilization question~\citep{wang2026sliderquant}.
Recent low-rank residual methods further refine where error should be
preserved or reconstructed~\citep{cho2026srr,park2026serq,li2026rescomp}.
We keep data-free frozen, calibrated no-backward, backward-assisted, and
globally trained methods in separate lanes because these are different
experimental interventions.

\begin{table}[t]
  \caption{Training-dependence lanes used throughout the comparison.}
  \label{tab:lane-summary}
  \centering
  \small
  \begin{tabular}{@{}clp{4.2cm}@{}}
    \toprule
    Lane & Backward? & Intervention \\
    \midrule
    A0 & No & Data-free frozen compression \\
    A1 & No & Calibration-only frozen compression \\
    B  & Yes & Encoder-side gradient/Hessian--vector products (HVPs), no global recovery \\
    C  & Yes & Learned or globally fine-tuned recovery \\
    \bottomrule
  \end{tabular}
\end{table}

The selected primary-source protocol and payload audit is reported in
\cref{tab:lanes}.  It is a literature taxonomy, not a protocol-matched
empirical leaderboard.
