\section{Theory and Diagnostic Predictions}
\label{sec:theory}

\paragraph{Curvature is not automatically a geometry.}
Let $\mathcal{L}(w)$ be held-out loss, $g=\nabla\mathcal{L}(w_0)$, and
$J_0=\nabla^2\mathcal{L}(w_0)$.  The task Hessian $J_0$ may be
indefinite, so $x^\top J_0x$ is not a squared norm and a ``Hessian
cosine'' based on it need not be an angle.  We reserve $H\succeq0$ for
a declared positive-semidefinite metric: an empirical Fisher, a
generalized Gauss--Newton matrix, the activation Gram proxy used in our
experiments, or $J_0+\lambda I$ only when $\lambda$ is large enough to
make it PSD.  We write
$\langle a,b\rangle_H=a^\top Hb$ and
$\|a\|_H^2=a^\top Ha$; zero-energy directions are excluded when a
correlation is reported.

\paragraph{Self, cross, and path terms.}
For sequential compression stages, define the realized increments
$\delta_i=w_i-w_{i-1}$ and $\delta=\sum_i\delta_i$.  This telescoping
identity is exact, although each $\delta_i$ depends on stage order.  A
local expansion around $w_0$ can be written
\begin{equation}
\label{eq:loss-decomposition}
\begin{split}
\mathcal{L}(w_0+\delta)-\mathcal{L}(w_0)
={}&g^\top\delta\\
&+\frac{1}{2}\sum_i q_i+\sum_{i<j}c_{ij}
+\varepsilon_H(\delta),\\
q_i={}&\|\delta_i\|_H^2,\\
c_{ij}={}&\langle\delta_i,\delta_j\rangle_H .
\end{split}
\end{equation}
Here $\varepsilon_H$ contains both the mismatch between $H$ and $J_0$
and curvature variation along the path.  If $H=J_0$ and the Hessian is
$M$-Lipschitz in the relevant trust region, then
$|\varepsilon_H(\delta)|\le M\|\delta\|_2^3/6$.
Thus a small quadratic cost alone does not justify dropping the
first-order term or evaluating only an interpolation point with
$\epsilon<1$.

\paragraph{What near-orthogonality guarantees.}
For $q_iq_j>0$, define
\begin{equation}
\rho_H(i,j)=\frac{c_{ij}}{\sqrt{q_iq_j}},\qquad
\mu=\max_{i\ne j}|\rho_H(i,j)|.
\end{equation}
For $k$ components in a common PSD metric,
\begin{equation}
\label{eq:coherence-bound}
\bigl(1-(k-1)\mu\bigr)\sum_iq_i
\le \Bigl\|\sum_i\delta_i\Bigr\|_H^2
\le \bigl(1+(k-1)\mu\bigr)\sum_iq_i,
\end{equation}
with a nontrivial lower bound when $(k-1)\mu<1$.
Consequently, near-zero $\rho_H(S,L)$ says that the sparse and low-rank
repair costs are approximately additive.  It does \emph{not} say that
their self costs are small, that either is orthogonal to quantization,
or that the combination beats a single codec.  Likewise,
$\rho_H<0$ is cancellation, not orthogonality.  Achieving prescribed
orthogonality further requires the repair degrees of freedom to span
the relevant $H$-moments and to remain feasible after coefficient
rounding and serialization (Appendix~\ref{app:theory-proofs}).

\paragraph{Comfort zones at a common rate.}
Let $s_i\ge0$ denote compression burden assigned to mechanism $i$ in a
continuous relaxation, and let $d_i(s_i)$ be its local loss cost after
its fixed codec charges have been paid.  If the $d_i$ are differentiable
and convex inside their comfort zones, an interior allocation solving
$\min\sum_i d_i(s_i)$ subject to $\sum_i s_i=S$ satisfies
\begin{equation}
\label{eq:kkt-comfort}
d_i'(s_i)=\lambda\quad(s_i>0),
\qquad d_i'(0^+)\ge\lambda\quad(s_i=0).
\end{equation}
This separable condition assumes that allocation-dependent interactions
are negligible or absorbed into the $d_i$; otherwise their partial
derivatives enter the KKT condition.
In particular, if a single mechanism $A$ carries all burden $S$ and
$d_A'(S)>d_B'(0^+)$, shifting a sufficiently small amount to $B$
reduces the relaxed cost.  This formalizes the comfort-zone intuition,
but ranks only local candidates: ranks, sparse groups, descriptors, and
alignment make the deployable problem discrete.

For a strict byte test, let $A$ be a realized reference endpoint at budget
$B$, $u$ the degradation needed to free bytes, and $r$ a repair.  A frontier
claim additionally requires $A$ to be the best eligible single-mechanism
endpoint under the same protocol.
For the realized candidate $C=A+u+r$, define
\begin{align}
P_A(u)&=\mathcal{L}(A+u)-\mathcal{L}(A),\nonumber\\
G_A(r)&=\mathcal{L}(A)-\mathcal{L}(A+r),\nonumber\\
\Gamma_A(u,r)&=\mathcal{L}(A+u+r)-\mathcal{L}(A+u)\nonumber\\
&\quad-\mathcal{L}(A+r)+\mathcal{L}(A).
\end{align}
Then the combination is better at the same rate whenever its actual
serialized length obeys $\mathcal{B}(C)\le\mathcal{B}(A)$ and
\begin{equation}
\label{eq:exact-exchange}
G_A(r)>P_A(u)+\Gamma_A(u,r).
\end{equation}
Locally, $\Gamma_A(u,r)$ contains the curvature cross term
$u^\top H_A r$ plus proxy and path errors, where $H_A$ is the declared
PSD proxy at $A$.  Equation~\eqref{eq:exact-exchange}, not pairwise
orthogonality, is the relevant sufficient condition.

\paragraph{Repair efficiency must charge the file.}
For a positive marginal file cost
$m_r=\mathcal{B}(A+r)-\mathcal{B}(A)$, we report
$\eta_{\mathrm{NLL}}=G_A(r)/m_r$.  The byte function includes codes,
scales, sparse values, indices and row pointers, low-rank factors,
descriptors, manifests, headers, and alignment/padding.  Parameter count or
value-only bits cannot substitute for this ledger.  If OBS value
re-estimation or a folded scale changes no final bytes, its direct
held-out recovery is reported separately rather than as infinite
efficiency; the zero-byte claim also requires decoding to contain no
hidden state.  A local quadratic ranking remains diagnostic until the
final decoded endpoint is evaluated on held-out data.

These distinctions first appeared in one seed on six selected Pythia-70M
MLP tensors: $S$ and $L$ are nearly orthogonal in the PSD proxy, both
correlate negatively with $Q$, and a conservative QSL candidate is
nevertheless worse than Q+L after tail padding to the same final bytes.  Three
additional, unpooled scalability-smoke jobs execute
the same diagnostic over full Pythia MLP scope and depth-stratified OPT and
Qwen MLP scopes.  They repeat the near-orthogonal S--L category and positive
strict-QSL-minus-QL endpoint sign for similarly underfilled candidates
(\cref{tab:scaling-pilot}), but they are
still single-seed observations with unequal scopes.  Multi-seed, common-scope,
multi-rate, and production-format tests remain confirmatory requirements.
